---
title: "simpleRegress"
author: "Jarrod Harris"
date: "Summer 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH ~ NECK, data = bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
bAGE <- lm(LENGTH ~ AGE, data = bears)
plot(LENGTH~AGE,data=bears)
abline(bAGE)
summary.lm(bAGE)
```

we want to compare the model with age to the other model with neck.
Ordering by adjusted R-squared in the model with neck we got 0.7476 and with the age model it was 0.5073. Using this method neck is better than age.

```{r}
bHEADLEN <- lm(LENGTH ~ HEADLEN, data = bears)
plot(LENGTH~HEADLEN,data=bears)
abline(bHEADLEN)
summary.lm(bHEADLEN)
```

 after comparing the neck model, the age model and the HEADLEN model. In order from biggest to the smallest  the difference is for HEADLEN  0.8434 , the neck model is  0.7476 , and lastly the age model is 0.5073. 
 
```{r}
bHEADWTH <- lm(LENGTH ~ HEADWTH, data = bears)
plot(LENGTH~HEADWTH,data=bears)
abline(bHEADWTH)
summary.lm(bHEADWTH)
```

After comparing the neck model, HEADLEN model, AGE, and HEADWTH in order from biggest to smallest and the results are HEADLEN-0.8434 , the neck model-0.7476, and lastly the HEADWTH model- 0.5321, lastly the AGE model-0.5073

```{r}
bCHEST <- lm(LENGTH ~ CHEST, data = bears)
plot(LENGTH~CHEST,data=bears)
abline(bCHEST)
summary.lm(bCHEST)
```
After comparing the neck model, HEADLEN model, AGE, HEADWTH, and CHEST model in order from biggest to smallest and the results with the HEADLEN model being the highest still at-0.8434 ,than the CHEST model being at- 0.7869,the neck model-0.7476, and lastly the HEADWTH model- 0.5321, lastly the AGE model-0.5073


```{r}
bWEIGHT <- lm(LENGTH ~ WEIGHT, data = bears)
plot(LENGTH~WEIGHT,data=bears)
abline(bWEIGHT)
summary.lm(bWEIGHT)
```

After comparing the neck model, HEADLEN model, AGE, HEADWTH, CHEST model, AND WEIGHT model in order from biggest to smallest and the results with the HEADLEN model being the highest still at-0.8434 ,than the CHEST model being at- 0.7869,the neck model-0.7476,the WEIGHT model-0.7423, the HEADWTH model- 0.5321,lastly the AGE model-0.5073
