---
title: "simpleRegress"
author: "Jarrod Harris"
date: "Summer 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH ~ NECK, data = bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
bAGE <- lm(LENGTH ~ AGE, data = bears)
plot(LENGTH~AGE,data=bears)
abline(bAGE)
summary.lm(bAGE)
```

we want to compare the model with age to the other model with neck.
Ordering by adjusted R-squared in the model with neck we got 0.7476 and with the age model it was 0.5073. Using this method neck is better than age.

```{r}
bHEADLEN <- lm(LENGTH ~ HEADLEN, data = bears)
plot(LENGTH~HEADLEN,data=bears)
abline(bHEADLEN)
summary.lm(bHEADLEN)
```

 after comparing the neck model, the age model and the HEADLEN model. In order from biggest to the smallest  the difference is for HEADLEN  0.8434 , the neck model is  0.7476 , and lastly the age model is 0.5073. 
 
 
```{r}
bHEADWTH <- lm(LENGTH ~ HEADWTH, data = bears)
plot(LENGTH~HEADWTH,data=bears)
abline(bHEADWTH)
summary.lm(bHEADWTH)
```